A Study of Residual Adapters for Multi-Domain Neural Machine Translation

### 2020-09-21 2020/wmt_20/mdadapt.tex
pretrained init scores, minor corrections

@@

Training gated residual adapters thus comprises three steps, instead of two for the baseline version:
\begin{enumerate}
\item learn a generic model with mixed corpora from multiple domains.
\item train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen. This model computes the posterior domain probability $P(k|h^L[t])$ for each word $w_t$, based on the representation computed by the last layer.
\item train the parameters of adapters with in-domain data separately for each domain, while freezing all the other parameters.
\end{enumerate}
\fyFuture{is this classifier important, can we train with the rest of the system ? Yes we could, but do we want the encoder to be good at classification?}\fyFuture{Check writing to indicate label smoothing for this has been used.}\fyFuture{Can we train a more complex training model ? should we us emore monolingual data ? etc}

@@

\begin{table*}[t!]
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Mixed}        & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\
    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
    \hline 
    \system{DC-Tag}      & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5  & 50.1 \\
    \system{DC-Feat}     & 37.7 & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 & 49.9  \\
    \system{LDR}            & 37.0  & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & 49.8    \\
    \system{TTM}           & 37.3  & 54.9 & 49.5 & 32.9 & 43.6 & 79.9 & 49.7   \\
    \system{DM}            & 35.6  & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4   \\ 
    \system{ADM}          & 36.4  & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5   \\
    \hline
    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
    \system{Res-Adap-MT}  & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
    %  \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
    \hline
  \end{tabular}
  \caption{Baseline translation performance of various multi-domain MT systems (En$\rightarrow$Fr).}
  \label{tab:performance-multi}
\end{table*}
The two variants of the residual adapter model included in this first round of experiment have been presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapters are jointly learned from scratch.

@@

\begin{table*}[htbp]
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
    \system{Mixed}          & 29.3 & 30.5 & 17.6 & 38.1 & 47.9 & 20.9  & 30.6 & 213M/0M\\
   \system{Res-Adap}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M \\ \hline
    \system{Res-Adap$_{(2,4,6)}$} & 29.7  & 30.5 & 18.8 & 49.6 & 47.1 & 20.6 &  32.7 & 213M/48M \\ \hline
    \system{Res-Adap$_{(6)}$}      & 29.5 & 30.4 & 18.1 & 49.1 & 46.9 & 20.4 & 32.4 & 213M/16M \\
   \system{Res-Adap$_{(4)}$}       & 29.7 & 30.4 & 18.1 & 49.6 & 47.0 & 20.6 & 32.6 & 213/16M\\
   \system{Res-Adap$_{(2)}$}       & 29.6 & 30.4 & 18.3 & 49.4 & 46.7 & 20.6 & 32.5  & 213M/16M\\
    \system{Res-Adap-WD}         & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
    \system{Res-Adap-LR}           & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1  & 213M/96M\\
    \hline
  \end{tabular}
  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$De). We report BLEU scores for each domain, as well as averages across domains. In column \domain{params}, we report the number of domain-agnostic/domain-specific parameters}
  \label{tab:performance-en-de-pos-reg}
\end{table*}

@@

\begin{table*}[htbp]
  \centering
  \begin{tabular}{|p{3cm}|*{7}{r|}|r|r|} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{rnd}} & \multicolumn{1}{c|}{\domain{news}} \\ \hline % & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline  
    \system{Mixed}             & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5     &  49.4 & 49.4 & 23.5 \\ %& 34s \\ %
    \system{FT-Full}             & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 53.8 & 32.5 & 20.2  \\ %& 34s \\
    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2   & 53.3 & 38.4 & 20.5 \\ %& 22s\\ 
    \system{Res-Adap-HW}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91.0  & 53.1 & 36.6 & 20.2 \\ %& 19s \\
    \system{Res-Adap-Gate}  & 38.0 & 57.5& 53.0 & 33.5 & 46.0 & 90.1  & 53.0 & 49.0 & 22.5 \\ %& 21s \\
%    \system{FT-Full-UEW}      & 37.9 & 56.0 & 52.1 & 33.7 & 44.9 & 89.1 &	52.3 & 46.5 & 22.11 \\ %& 34s \\
    \hline
  \end{tabular}
  \caption{Translation performance of highway and gated variants for En$\rightarrow$Fr.
    % Column \domain{PROC} is time (in seconds) of 100 iterations with batch size 12,288 tokens
  }
  \label{tab:performance-random}
\end{table*}
\fyFuture{Computation time should also report the time it takes for the complete tuning process, not one iteration which will be more or less the same}.
\section{Related Work \label{sec:related}}
\mpDone{related work}

>>>>

\usepackage{footnote}
\makesavenoteenv{tabular*}
\makesavenoteenv{table*}

@@

Training gated residual adapters thus comprises three steps, instead of two for the baseline version:
\begin{enumerate}
\item learn a generic model with mixed corpora from multiple domains.
\item train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen. This model computes the posterior domain probability $P(k|h^L[t])$ for each word $w_t$, based on the representation computed by the last layer.
\item train the parameters of adapters with in-domain data separately for each domain, while freezing all the other parameters.
\end{enumerate}
\fyFuture{is this classifier important, can we train with the rest of the system ? Yes we could, but do we want the encoder to be good at classification?}\fyFuture{Check writing to indicate label smoothing for this has been used.}\fyFuture{Can we train a more complex training model ? should we use more monolingual data ? etc}

@@

\begin{table*}[t!]
  \centering
  \begin{tabular}{|p{4.3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Mixed}        & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\
    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
    \hline 
    \system{DC-Tag}      & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5  & 50.1 \\
    \system{DC-Feat}     & 37.7 & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 & 49.9  \\
    \system{LDR}            & 37.0  & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & 49.8    \\
    \system{TTM}           & 37.3  & 54.9 & 49.5 & 32.9 & 43.6 & 79.9 & 49.7   \\
    \system{DM}            & 35.6  & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4   \\ 
    \system{ADM}          & 36.4  & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5   \\
    \hline
    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
    \system{Res-Adap-MT}  & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
    \system{Res-Adap-MT-ptr\footnote{We initialized shared parameters by values of parameters of \system{mixed} model and train Res-Adap using multi-task learning}} & 37.47 & 57.06 & 52.39 & 33.72 & 46.17 & 89.47 & 52.7 \\
    %  \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
    \hline
  \end{tabular}
  \caption{Baseline translation performance of various multi-domain MT systems (En$\rightarrow$Fr).}
  \label{tab:performance-multi}
\end{table*}
The two variants of the residual adapter model included in this first round of experiment have been presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapters are jointly learned from scratch.

@@

\begin{table*}[htbp]
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
    \system{Mixed}          & 29.3 & 30.5 & 17.6 & 38.1 & 47.9 & 20.9  & 30.6 & 213M/0M\\
   \system{Res-Adap}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M \\ \hline
    \system{Res-Adap$_{(2,4,6)}$} & 29.7  & 30.5 & 18.8 & 49.6 & 47.1 & 20.6 &  32.7 & 213M/48M \\ 
    \system{Res-Adap$_{(6)}$}      & 29.5 & 30.4 & 18.1 & 49.1 & 46.9 & 20.4 & 32.4 & 213M/16M \\
   \system{Res-Adap$_{(4)}$}       & 29.7 & 30.4 & 18.1 & 49.6 & 47.0 & 20.6 & 32.6 & 213M/16M\\
   \system{Res-Adap$_{(2)}$}       & 29.6 & 30.4 & 18.3 & 49.4 & 46.7 & 20.6 & 32.5  & 213M/16M\\
   \hline
    \system{Res-Adap-WD}         & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
    \system{Res-Adap-LR}           & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1  & 213M/96M\\
    \hline
  \end{tabular}
  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$De). We report BLEU scores for each domain, as well as averages across domains. In column \domain{params}, we report the number of domain-agnostic/domain-specific parameters}
  \label{tab:performance-en-de-pos-reg}
\end{table*}

@@

\begin{table*}[htbp]
  \centering
  \begin{tabular}{|p{4.8cm}|*{7}{r|}|r|r|} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{rnd}} & \multicolumn{1}{c|}{\domain{news}} \\ \hline % & \multicolumn{1}{c|}{\domain{PROC}} \\ \hline  
    \system{Mixed}             & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5     &  49.4 & 49.4 & 23.5 \\ %& 34s \\ %
    \system{FT-Full}             & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8   & 53.8 & 32.5 & 20.2  \\ %& 34s \\
    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2   & 53.3 & 38.4 & 20.5 \\ %& 22s\\ 
    \system{Res-Adap-HW}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91.0  & 53.1 & 36.6 & 20.2 \\ %& 19s \\
    \system{Res-Adap-HW-MT} & 37.4 & 56.4 & 52.1 & 33.7 & 44.8 & 89.8 & 52.4 & 27.1 & 20.4 \\ %&  \\
    \system{Res-Adap-HW-MT-ptr \footnote{We initialized shared parameters by values of parameters of \system{mixed} model and train Res-Adap-HW using multi-task learning}} & 37.7 & 57 & 52.5 & 33.5 & 46.1 & 89 & 52.6 & 46.5 & 21.4 \\ %&  \\
    \system{Res-Adap-Gate}  & 38.0 & 57.5& 53.0 & 33.5 & 46.0 & 90.1  & 53.0 & 49.0 & 22.5 \\ %& 21s \\
%    \system{FT-Full-UEW}      & 37.9 & 56.0 & 52.1 & 33.7 & 44.9 & 89.1 &	52.3 & 46.5 & 22.11 \\ %& 34s \\
    \hline
  \end{tabular}
  \caption{Translation performance of highway and gated variants for En$\rightarrow$Fr.
    % Column \domain{PROC} is time (in seconds) of 100 iterations with batch size 12,288 tokens
  }
  \label{tab:performance-random}
\end{table*}
\fyFuture{Computation time should also report the time it takes for the complete tuning process, not one iteration which will be more or less the same}.
\section{Related Work \label{sec:related}}
\mpDone{related work}
